---
title: "testis-BulkDeconv"
author: "BH"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, readxl, ggplot2, ggsci, ggthemes, vroom, readr, tidyverse, DWLS, stringr, Metrics, ggbreak)

theme_set(theme_few() + theme(text = element_text(size=28, family="Myriad Pro")))

```

Load in each methods combined psuedobulk (all magics, all alras, etc.). Then run dWLS deconv on those and remove the same-same comparisons. Only use the D7_diff files


Load the signatures matrices
```{r sigs}

swd <- list.dirs("~/OneDrive - UW/data for deconv/Signature matrices impute+no impute (D)/Testis/cross_compare_sigs", full.names = TRUE)

# Initialize an empty list to hold sublists for each type
all_matrices <- list()

type <- c("noIMP", "ALRA", "SAVER", "MAGIC")

for (dir in swd) {

    for (t in type) {
        if (grepl(t, dir)) {
        
            if (is.null(all_matrices[[t]])) {
                all_matrices[[t]] <- list()
            }
            
            rds_files <- list.files(dir, pattern = "Sig.rds", full.names = TRUE, recursive = TRUE)
            
            for (rds_file in rds_files) {
            
                data <- readRDS(rds_file)
                file_name_part <- basename(dirname(rds_file))
                new_name <- paste0(file_name_part)
                all_matrices[[t]][[new_name]] <- data   }} }}



```

```{r rnaseq}

testisrna <- readRDS("~/OneDrive - UW/data for deconv/all_testis_rnaseq.rds")

testis_export <- testisrna %>% rownames_to_column(var= 'geneid')

write.table(testis_export, "~/OneDrive - UW/data for deconv/testisrna.txt", sep = "\t", row.names = FALSE,col.names = TRUE, quote = FALSE)

```


FOR TESTIS WE WANT SVR-ALRA

```{r deconv draft cross, include=FALSE}

  DWLS <- data.frame()
  SVR <- data.frame()
  OLS <- data.frame()


runDWLSall <- function(df_in, sig_matrix_in) {
  rows <- intersect(rownames(sig_matrix_in), rownames(df_in))
  
  df <- df_in[rows, ]
  sig_matrix_in <- sig_matrix_in[rows, ]
  
  samples <- colnames(df)
  
  for (i in samples) {
    name <- i
    dataBulk <- as.numeric(df[, i])
    Signature <- as.matrix(sig_matrix_in)
    Signature <- 2^Signature - 1  
    
    solSVR <- solveSVR(Signature, dataBulk)

    SVR <- rbind(SVR, solSVR)
  }
  
  # Ensure correct column names and add method columns before binding results

  colnames(SVR) <- c(colnames(Signature))
  SVR$method <- 'SVR'
  

  out <- rbind(SVR)
  out$Mixture <- samples

  return(out)
}


noimp_allcross <- lapply(all_matrices$ALRA, runDWLSall, df_in = testisrna)

result <- lapply(names(noimp_allcross), function(name) {
  df <- noimp_allcross[[name]]
  
  if ("StromaLeydig" %in% colnames(df)) {
    df <- df %>%
      rename(Leydig = StromaLeydig)  }
  
  df <- df %>%
    select(Sertoli, Leydig, method, PTM, Mixture) %>% mutate(sig_ref = name)  
  return(df)
})

ALRA_all <- bind_rows(result)
ALRA_all$imp <- 'ALRA'
ALRA_all$title <- str_extract(ALRA_all$sig_ref, "^.*(?=_[^_]+$)")

ALRA_all <- ALRA_all %>% select(-sig_ref)

```

load in the cibersorts

```{r cibersorts load}
fp <- ("~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/bulkDeconv/cibersort")
files <- list.files(path = fp, pattern = "*.csv", full.names = TRUE)

cb_testis <- list()

for (file in files) {
  data <- read_csv(file)

  basename_file <- tools::file_path_sans_ext(basename(file))
  method_value <- toupper(str_extract(basename_file, "(?<=_)[^_]+(?=$)"))
 samp <- str_extract(basename_file, "^.*(?=_[^_]+$)")
  
  data <- data %>%
    { if ("Stroma/Leydig" %in% colnames(.)) rename(., Leydig = `Stroma/Leydig`) else . } %>%
    select(Mixture, Leydig, Sertoli, PTM) %>%
    mutate(name = basename_file,
      imp = method_value,
      title = samp)
  
  
  cb_testis[[basename_file]] <- data
}

cb_testis <- bind_rows(cb_testis)%>% select(-name)%>% mutate(method='CIBERSORTx') %>% mutate(
                                                             ref = recode(title,
                                                                    'm_neo_d2' = 'm_i_2',
                                                                    'm_d7_DIFF' = 'm_i_7'))%>% select(-title)


ALRA_all <-  ALRA_all %>% mutate(ref = recode(title,
                                  'm_neo_d2' = 'm_i_2',
                                  'm_d7_DIFF' = 'm_i_7'))%>% 
               filter(ref != "m_n_d7_SL")%>% select(-title)
```

combine them

```{r combine}
testis_metadata <- read_xlsx('~/OneDrive - UW/data for deconv/rna-seq data (A)/testis_metadata.xlsx')


testis_deconv <- rbind(ALRA_all, cb_testis)%>% pivot_longer(cols = c('Sertoli','Leydig','PTM'), names_to = 'celltype')%>%mutate(id = paste0(method,'-',imp))%>% left_join(testis_metadata, by= 'Mixture')

treated <- c('H01','H02','H03')

testis_deconv <-  testis_deconv %>% mutate(ref = recode(ref,
                                  'h_a_112' = 'Adult-H',
                                  'h_i_120' = 'Infant-120',
                                  'h_i_124' = 'Infant-124',
                                  'm_i_2' = 'Day 7-M',
                                  'm_i_7' = 'Day 2-M'))%>%
                      filter(!Mixture %in% treated)



```

plotting
```{r plotting}
plots <- ggplot(data=testis_deconv%>%filter(exp=='baseline'), aes(x=factor(time), y=value, fill = celltype)) +
  geom_boxplot(alpha = 0.45, outliers = FALSE) + 
geom_jitter(aes(x=factor(time), y=value, color = celltype, shape = ref), 
              position = position_dodge(width = 0.75), 
              size = 3, alpha = 0.8) +  
  scale_y_continuous(name="Cell Proportion", breaks = c(0, 0.25, 0.5, 0.75, 1.0), limits = c(0,1))+
  scale_x_discrete(labels=c("0" = "Tissue", 
                              "3" = "Day 3",
                              "5" = "Day 5",
                              "7" = "Day 7"))+
  scale_fill_d3()+
  scale_color_d3()+
  labs(shape="Reference", color= "Cell Type")+
  xlab("Culture Time")+
  guides(fill = "none")+
  theme(legend.position="right", axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~id)


plots

ggsave(plots, dpi = 600, filename = "~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/plots/testis-baseline-deconv.png", width = 16, height=6, units = 'in')

```



plot the LH ones:
```{r plotting LH}
plots2 <- ggplot(data=testis_deconv%>%filter(exp=='lh'), aes(x=factor(time), y=value, fill = celltype)) +
  geom_boxplot(alpha = 0.45, outliers = FALSE) + 
geom_jitter(aes(x=factor(time), y=value, color = celltype, shape = ref), 
              position = position_dodge(width = 0.75), 
              size = 3, alpha = 0.8) +  
  scale_y_continuous(name="Cell Proportion", breaks = c(0, 0.25, 0.5, 0.75, 1.0), limits = c(0,1))+
    scale_x_discrete(labels=c("0" = "Day 10 Baseline", 
                              "2" = "2 hours",
                              "12" = "12 hours",
                              "48" = "48 hours"))+
  scale_fill_d3()+
  scale_color_d3()+
  labs(shape="Reference", color= "Cell Type")+
  xlab("Hours after LH Stimulus")+
  guides(fill = "none")+
  theme(legend.position="right", axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~id)


plots2

ggsave(plots2, dpi = 600, filename = "~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/plots/testis-LH-deconv.png", width = 16, height=6, units = 'in')

```


Germ Cells:

human adult: d_spg and SSCs
h120: Germ
h124: Germ
md7: diff_SPG and ProSGSSC
md2: ProSGSSC

```{r germ cell plots}

germs <- c("d_spg", "SSCs", "Germ", "diff_SPG", "ProSGSSC", "meiosis_spcs")

pull_germ <- function(df, ref_name) {
  df %>%
    select(any_of(germs), method, Mixture) %>%
    pivot_longer(cols = any_of(germs), names_to = "type", values_to = "percent") %>%
    mutate(ref = ref_name) 
}

# Assuming noimp_allcross is a list of data frames
dfs_long1 <- map2(noimp_allcross, names(noimp_allcross), pull_germ)


  df_combined <- bind_rows(dfs_long1)%>% filter(!Mixture %in% treated)%>% mutate(imp = "ALRA")

  # samp <- str_extract(basename_file, "^.*(?=_[^_]+$)")

  
  
all_germ <- function(df,ref_name) {
  df %>%
    select(any_of(germs), method, Mixture) %>%
    mutate(total = rowSums(select(., -method, -Mixture), na.rm = TRUE)) %>%
    mutate(ref = ref_name) 
   
}

dfs_long2 <- map2(noimp_allcross, names(noimp_allcross), all_germ)
df_same <- bind_rows(dfs_long2)%>% filter(!Mixture %in% treated)%>% mutate(imp = "ALRA")
df_same$title <- str_extract(df_same$ref, "^.*(?=_[^_]+$)")
df_same <- df_same %>% mutate(sigref = recode(title,
                              'm_neo_d2' = 'm_i_2',
                                'm_d7_DIFF' = 'm_i_7'))%>%
  filter(sigref != "m_n_d7_SL")%>% select(-c(title,ref)) 

```


```{r cibersorts load}
fp <- ("~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/bulkDeconv/cibersort")
files <- list.files(path = fp, pattern = "*.csv", full.names = TRUE)

cb_germ <- list()

for (file in files) {
  data <- read_csv(file)

  basename_file <- tools::file_path_sans_ext(basename(file))
  method_value <- toupper(str_extract(basename_file, "(?<=_)[^_]+(?=$)"))
  samp <- str_extract(basename_file, "^.*(?=_[^_]+$)")
  
  data <- data %>%
    mutate(name = basename_file,
      imp = method_value,
      title = samp)
  
  cb_germ[[basename_file]] <- data
}

cb_germ <- bind_rows(cb_germ)%>% select(c(Mixture,d_spg, SSCs, meiosis_spcs, imp, Germ, `Pro-SG/SSC`, diff_SPG, title)) %>% mutate(method='CIBERSORTx')%>%rename(ProSGSSC =`Pro-SG/SSC`)%>%
   mutate(total = rowSums(select(., -method, -Mixture, -imp, -title), na.rm = TRUE))


cb_germ <-  cb_germ %>% mutate(sigref = recode(title,
                              'm_neo_d2' = 'm_i_2',
                                'm_d7_DIFF' = 'm_i_7'))%>% 
              filter(sigref != "m_n_d7_SL") %>%select(-title)
```


combine the germ cell ones
```{r all germs}

germ_plot <- rbind(cb_germ, df_same)%>% left_join(testis_metadata, by= 'Mixture')%>%mutate(id = paste0(method,'-',imp)) %>% mutate(sigref = recode(sigref,
                                  'h_a_112' = 'Adult-H',
                                  'h_i_120' = 'Infant-120',
                                  'h_i_124' = 'Infant-124',
                                  'm_i_2' = 'Day 7-M',
                                  'm_i_7' = 'Day 2-M'))%>%
                      filter(!Mixture %in% treated)

```


plot the germ cell ones
```{r plotting baseline}
plots4 <- ggplot(data=germ_plot%>%filter(exp=='baseline'), aes(x=factor(time), y=total)) +
  geom_boxplot(alpha = 0.45, outliers = FALSE, fill="gray") + 
geom_point(aes(x=factor(time), y=total, color = sigref), 
              position = position_dodge(width = 0.1), 
              size = 3, alpha = 0.8) +  
  scale_y_continuous(name="Cell Type", breaks = c(0, 0.25, 0.5, 0.75, 1.0), limits = c(0,1))+
  scale_x_discrete(labels=c("0" = "Tissue", 
                              "3" = "Day 3",
                              "5" = "Day 5",
                              "7" = "Day 7"))+
  scale_fill_d3()+
  scale_color_d3()+
  labs(shape="Reference", color= "Cell Type")+
  xlab("Culture Time")+
  guides(fill = "none")+
  theme(legend.position="right", axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~id)


plots4

ggsave(plots4, dpi = 600, filename = "~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/plots/testis-baseline_GERM.png", width = 16, height=5, units = 'in')

```




```{r plotting LH}
plots5 <- ggplot(data=germ_plot%>%filter(exp=='lh'), aes(x=factor(time), y=total)) +
  geom_boxplot(alpha = 0.45, outliers = FALSE, fill="gray") + 
geom_point(aes(x=factor(time), y=total, color = sigref), 
              position = position_dodge(width = 0.1), 
              size = 3, alpha = 0.8) +  
  scale_y_continuous(name="Cell Type %", breaks = c(0, 0.25, 0.5, 0.75, 1.0), limits = c(0,1))+
  scale_fill_d3()+
  scale_color_d3()+
  labs(shape="Reference", color= "Cell Type")+
  xlab("Hours after LH Treatment")+
  guides(fill = "none")+
  theme(legend.position="right", axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~id)


plots5

ggsave(plots5, dpi = 600, filename = "~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/plots/testis-LH-GERM.png", width = 16, height=5, units = 'in')

```


```{r plotting LH}

germ_plot$lhtime <- factor(germ_plot$lhtime, levels = c("0","3","5","7","10","12"))

plots_dates <- ggplot(data=germ_plot%>%filter(lhtime!='na'), aes(x=factor(lhtime), y=total)) +
  geom_boxplot(alpha = 0.45, outliers = FALSE, fill="gray") + 
geom_point(aes(x=factor(lhtime), y=total, color = sigref), 
              position = position_dodge(width = 0.1), 
              size = 3, alpha = 0.8) +  
  scale_y_continuous(name="Cell Proportion", breaks = c(0, 0.25, 0.5, 0.75, 1.0), limits = c(0,1))+
  scale_x_discrete(labels=c("0" = "Tissue", 
                              "3" = "Day 3",
                              "5" = "Day 5",
                              "7" = "Day 7",
                              "10" = "Day 10",
                              "12" = "Day 12"))+
  scale_fill_d3()+
  scale_color_d3()+
  labs(shape="Reference", color= "Reference")+
  xlab("Culture Day")+
  guides(fill = "none")+
  theme(legend.position="right", axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~id)


plots_dates

ggsave(plots_dates, dpi = 600, filename = "~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/plots/testis-LHallTIME-GERM.png", width = 16, height=5, units = 'in')

```
