---
title: "testis-crosscompare_DWLS"
author: "BH"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, readxl, ggplot2, ggsci, ggthemes, vroom, readr, tidyverse, DWLS, stringr, Metrics)

theme_set(theme_few() + theme(text = element_text(size=28, family="Myriad Pro")))

```

load objects
```{r objects}

human_to_rat <- readRDS("~/OneDrive - UW/data for deconv/public data (B)/Testis/Rat_to_Human.rds")
# Keep only the Gene Names
human_to_rat <- human_to_rat[,c("Human_GRCh38_GeneSymbol","Rat_mRatBN7.2_GeneSymbol")]
colnames(human_to_rat) <- c("human","rat")
human_to_rat <- human_to_rat %>% distinct()

Rat_to_Mouse <- readRDS("~/OneDrive - UW/data for deconv/public data (B)/Testis/Rat_to_Mouse.rds")
# Keep only the Gene Names
Rat_to_Mouse <- Rat_to_Mouse[,c("Mouse_GRCm39_GeneSymbol","Rat_mRatBN7.2_GeneSymbol")]
colnames(Rat_to_Mouse) <- c("mouse","rat")
Rat_to_Mouse <- Rat_to_Mouse %>% distinct()

files_out <- "~/OneDrive - UW/data for deconv/pseudobulks (E)/Testis/combined/"

```
Load in each methods combined psuedobulk (all magics, all alras, etc.). Then run dWLS deconv on those and remove the same-same comparisons. Only use the D7_diff files

load and combine the noIMP PBs
```{r load noIMP PBs}

dfwd <- ("~/OneDrive - UW/data for deconv/pseudobulks (E)/Testis/")

# no Imp
noimps <- list.files(paste0(dfwd,"noImp"),  pattern = "*_combined.txt", full.names = TRUE)

noimp_PBs <- list()

for (file in noimps) {
  
  df <- read.delim(file)
  
    if (startsWith(basename(file), "m_")) {
    df <- df %>%  distinct(V1, .keep_all = TRUE) %>%  left_join(Rat_to_Mouse %>% distinct(mouse, .keep_all = TRUE), by = c('V1' = 'mouse')) %>%
  filter(!is.na(rat)) %>% distinct(rat, .keep_all = TRUE) %>% column_to_rownames("rat") %>%select(-V1)%>% rownames_to_column(var = "rowname")  
    noimp_PBs[[basename(file)]] <- df  
  } else {
    df <- left_join(df,human_to_rat, by = c('V1' = 'human')) %>% filter(!is.na(rat)) %>% column_to_rownames("rat") %>% select(-V1) %>% rownames_to_column(var = "rowname")  

    noimp_PBs[[basename(file)]] <- df  
  }
}

all_noimp_PBs <- Reduce(function(x, y) inner_join(x, y, by = "rowname"), noimp_PBs) %>% column_to_rownames(var = "rowname")

all_noimp_PBs <- all_noimp_PBs %>% rownames_to_column('gene')

write.table(all_noimp_PBs, file = paste0(files_out, "all_noimp_PBs.txt"), sep = "\t", quote = FALSE, row.names = FALSE,col.names = TRUE)
```

load and combine the ALRA PBs
```{r load ALRA PBs}

dfwd <- ("~/OneDrive - UW/data for deconv/pseudobulks (E)/Testis/")

# no Imp
alra <- list.files(paste0(dfwd,"ALRA"),  pattern = "*_combined.txt", full.names = TRUE)

dfs <- list()

for (file in alra) {
  
  df <- read.delim(file)
  
    if (startsWith(basename(file), "m_")) {
    df <- df %>%  distinct(V1, .keep_all = TRUE) %>%  left_join(Rat_to_Mouse %>% distinct(mouse, .keep_all = TRUE), by = c('V1' = 'mouse')) %>%
  filter(!is.na(rat)) %>% distinct(rat, .keep_all = TRUE) %>% column_to_rownames("rat") %>%select(-V1)%>% rownames_to_column(var = "rowname")  
    dfs[[basename(file)]] <- df  
  } else {
    df <- left_join(df,human_to_rat, by = c('V1' = 'human')) %>% filter(!is.na(rat)) %>% column_to_rownames("rat") %>% select(-V1) %>% rownames_to_column(var = "rowname")  

    dfs[[basename(file)]] <- df  
  }
}

all_ALRA_PBs <- Reduce(function(x, y) inner_join(x, y, by = "rowname"), dfs) %>% column_to_rownames(var = "rowname")


all_ALRA_PBs <- all_ALRA_PBs %>% rownames_to_column('gene')

write.table(all_ALRA_PBs, file = paste0(files_out, "all_ALRA_PBs.txt"), sep = "\t", quote = FALSE, row.names = FALSE,col.names = TRUE)



```

load and combine the SAVER PBs
```{r load SAVER PBs}

dfwd <- ("~/OneDrive - UW/data for deconv/pseudobulks (E)/Testis/")

# no Imp
svr <- list.files(paste0(dfwd,"SAVER"),  pattern = "*_combined.txt", full.names = TRUE)

dfs <- list()

for (file in svr) {
  
  df <- read.delim(file)
  
    if (startsWith(basename(file), "m_")) {
    df <- df %>%  distinct(V1, .keep_all = TRUE) %>%  left_join(Rat_to_Mouse %>% distinct(mouse, .keep_all = TRUE), by = c('V1' = 'mouse')) %>%
  filter(!is.na(rat)) %>% distinct(rat, .keep_all = TRUE) %>% column_to_rownames("rat") %>%select(-V1)%>% rownames_to_column(var = "rowname")  
    dfs[[basename(file)]] <- df  
  } else {
    df <- left_join(df,human_to_rat, by = c('V1' = 'human')) %>% filter(!is.na(rat)) %>% column_to_rownames("rat") %>% select(-V1) %>% rownames_to_column(var = "rowname")  

    dfs[[basename(file)]] <- df  
  }
}

all_SAVER_PBs <- Reduce(function(x, y) inner_join(x, y, by = "rowname"), dfs) %>% column_to_rownames(var = "rowname")

all_SAVER_PBs <- all_SAVER_PBs %>% rownames_to_column('gene')

write.table(all_SAVER_PBs, file = paste0(files_out, "all_SAVER_PBs.txt"), sep = "\t", quote = FALSE, row.names = FALSE,col.names = TRUE)


```

load and combine the MAGIC PBs
```{r load MAGIC PBs}

dfwd <- ("~/OneDrive - UW/data for deconv/pseudobulks (E)/Testis/")

# no Imp
magic <- list.files(paste0(dfwd,"MAGIC"),  pattern = "*_combined.txt", full.names = TRUE)

dfs <- list()

for (file in magic) {
  
  df <- read.delim(file)
  
    if (startsWith(basename(file), "m")) {
    df <- df %>%  distinct(V1, .keep_all = TRUE) %>%  left_join(Rat_to_Mouse %>% distinct(mouse, .keep_all = TRUE), by = c('V1' = 'mouse')) %>%
  filter(!is.na(rat)) %>% distinct(rat, .keep_all = TRUE) %>% column_to_rownames("rat") %>%select(-V1)%>% rownames_to_column(var = "rowname")  
    dfs[[basename(file)]] <- df  
  } else {
    df <- left_join(df,human_to_rat, by = c('V1' = 'human')) %>% filter(!is.na(rat)) %>% column_to_rownames("rat") %>% select(-V1) %>% rownames_to_column(var = "rowname")  

    dfs[[basename(file)]] <- df  
  }
}

all_MAGIC_PBs <- Reduce(function(x, y) inner_join(x, y, by = "rowname"), dfs) %>% column_to_rownames(var = "rowname")

all_MAGIC_PBs <- all_MAGIC_PBs %>% rownames_to_column('gene')

write.table(all_MAGIC_PBs, file = paste0(files_out, "all_MAGIC_PBs.txt"), sep = "\t", quote = FALSE, row.names = FALSE,col.names = TRUE)
```

Load the signatures matrices
```{r sigs}

swd <- list.dirs("~/OneDrive - UW/data for deconv/Signature matrices impute+no impute (D)/Testis/cross_compare_sigs", full.names = TRUE)

# Initialize an empty list to hold sublists for each type
all_matrices <- list()

type <- c("noIMP", "ALRA", "SAVER", "MAGIC")

for (dir in swd) {

    for (t in type) {
        if (grepl(t, dir)) {
        
            if (is.null(all_matrices[[t]])) {
                all_matrices[[t]] <- list()
            }
            
            rds_files <- list.files(dir, pattern = "Sig.rds", full.names = TRUE, recursive = TRUE)
            
            for (rds_file in rds_files) {
            
                data <- readRDS(rds_file)
                file_name_part <- basename(dirname(rds_file))
                new_name <- paste0(t, "_", file_name_part, "_sig")
                all_matrices[[t]][[new_name]] <- data   }} }}



```



deconvolution draft code
```{r deconv draft cross, include=FALSE}

df_in <- all_noimp_PBs

  DWLS <- data.frame()
  SVR <- data.frame()
  OLS <- data.frame()


runDWLSall <- function(df_in, sig_matrix_in) {
  rows <- intersect(rownames(sig_matrix_in), rownames(df_in))
  
  df <- df_in[rows, ]
  sig_matrix_in <- sig_matrix_in[rows, ]
  
  samples <- colnames(df)
  
  for (i in samples) {
    name <- i
    dataBulk <- as.numeric(df[, i])
    Signature <- as.matrix(sig_matrix_in)
    Signature <- 2^Signature - 1  
    
    solDWLS <- solveDampenedWLS(Signature, dataBulk)
    solSVR <- solveSVR(Signature, dataBulk)
    solOLS <- solveOLS(Signature, dataBulk)
    
    DWLS <- rbind(DWLS, solDWLS)
    SVR <- rbind(SVR, solSVR)
    OLS <- rbind(OLS, solOLS)
  }
  
  # Ensure correct column names and add method columns before binding results
  colnames(DWLS) <- c(colnames(Signature))
  DWLS$method <- 'DWLS'

  colnames(SVR) <- c(colnames(Signature))
  SVR$method <- 'SVR'
  
  colnames(OLS) <- c(colnames(Signature))
  OLS$method <- 'OLS'

  out <- rbind(DWLS, OLS, SVR)
  out$ref <- rep(samples, times = 3)

  return(out)
}


noimp_allcross <- lapply(all_matrices$noIMP, runDWLSall, df_in = all_noimp_PBs)

result <- lapply(names(noimp_allcross), function(name) {
  df <- noimp_allcross[[name]]
  
  if ("StromaLeydig" %in% colnames(df)) {
    df <- df %>%
      rename(Leydig = StromaLeydig)  }
  
  df <- df %>%
    select(Sertoli, Leydig, method, ref) %>% mutate(list_item_name = name)  
  return(df)
})

noImp_cross <- bind_rows(result)
noImp_cross$imp <- 'none'

```

now the ALRAS
```{r alra cross compare, include=FALSE}
alra_allcross <- lapply(all_matrices$ALRA, runDWLSall, df_in = all_ALRA_PBs)

result <- lapply(names(alra_allcross), function(name) {
  df <- alra_allcross[[name]]
  
  if ("StromaLeydig" %in% colnames(df)) {
    df <- df %>%
      rename(Leydig = StromaLeydig)  }
  
  df <- df %>%
    select(Sertoli, Leydig, method,ref) %>% mutate(list_item_name = name)  
  return(df)
})

alra_cross <- bind_rows(result)
alra_cross$imp <- 'alra'
```

now the SAVERs
```{r saver cross compare, include=FALSE}
saver_allcross <- lapply(all_matrices$SAVER, runDWLSall, df_in = all_SAVER_PBs)

result <- lapply(names(saver_allcross), function(name) {
  df <- saver_allcross[[name]]
  
  if ("StromaLeydig" %in% colnames(df)) {
    df <- df %>%
      rename(Leydig = StromaLeydig)  }
  
  df <- df %>%
    select(Sertoli, Leydig, method, ref) %>% mutate(list_item_name = name)  
  return(df)
})

saver_cross <- bind_rows(result)
saver_cross$imp <- 'saver'
```


now the MAGICs
```{r magic cross compare, include=FALSE}
magic_allcross <- lapply(all_matrices$MAGIC, runDWLSall, df_in = all_MAGIC_PBs)

result <- lapply(names(magic_allcross), function(name) {
  df <- magic_allcross[[name]]
  
  if ("StromaLeydig" %in% colnames(df)) {
    df <- df %>%
      rename(Leydig = StromaLeydig)  }
  
  df <- df %>%
    select(Sertoli, Leydig, method, ref) %>% mutate(list_item_name = name)  
  return(df)
})

magic_cross <- bind_rows(result)
magic_cross$imp <- 'magic'
```


combine all the objects for plotting
```{r combine}

all_crosses <- rbind(magic_cross,alra_cross,saver_cross, noImp_cross) %>% 
  mutate(trues = str_extract(ref, "\\d{2}$"),
         PB = str_sub(ref, 1, nchar(ref) - 2))


all_crosses$PBbase<- sub("^(.*_[^_]+)_.*$", "\\1", all_crosses$PB)
all_crosses$PBbase <- sub("_[^_]*$", "", all_crosses$PBbase)

all_crosses$REFbase<- sub("^(.*_[^_]+)_.*$", "\\1", all_crosses$list_item_name)
all_crosses$REFbase <- sub("_[^_]*$", "", all_crosses$REFbase)
all_crosses$REFbase <- sub("^[^_]+_", "", all_crosses$REFbase)

saveRDS(all_crosses,"~/OneDrive - UW/data for deconv/Signature matrices impute+no impute (D)/Testis/cross_compare_sigs/all_crosses_wocb.rds")

```

```{r file load}
all_crosses <- readRDS("~/OneDrive - UW/data for deconv/Signature matrices impute+no impute (D)/Testis/cross_compare_sigs/all_crosses_wocb.rds")
cb_testis <- readRDS("~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/cibersort_cross_outputs/cb_testis.rds") %>% rename(trues = pb)%>% select(Mixture, Sertoli, Leydig, cell, PBbase, REFbase, trues, imp)%>%
  mutate(method = 'CIBERSORTx')

```

matching checks


table(all_crosses$PBbase)
#.   adult112   H_i_120   H_i_124   H_I.120   H_I.124   h_n_124    hi_124 m_d7_diff    m_n_d7 m_n_d7_d2 m_n_d7_SL  m_neo_d2  m_neo_d7     nh120
 #     360        90        90        90        90        90        90       360        72        18       180       360        90     180
       

table(all_crosses$REFbase)

 #   h_a_112   h_i_120   h_i_124 m_d7_diff m_d7_DIFF m_n_d7_SL  m_neo_d2  m_neo_d7 
  #    360       360       360        90       270       270       360        90 



FINAL LABELS --->   h_a_112     h_i_120     h_i_124     m_d7_diff     m_n_d7_SL    m_n_d2

```{r recode}

all_crosses <- all_crosses %>%
  mutate(REFbase = recode(REFbase,
                    "m_d7_DIFF" = "m_d7_diff",  
                    "m_n_d7" = "m_neo_d7",
                   "adult112" = "h_a_112", 
                    
                    "H_i_120" = "h_i_120", 
                    "H_I.120" = "h_i_120", 
                    "nh120" = "h_i_120", 
                    
                    "H_i_124" = "h_i_124", 
                    "H_I.124" = "h_i_124", 
                    "h_n_124" = "h_i_124",
                    "hi_124" = "h_i_124",
                    

                    "m_neo_d7" = "m_n_d7_SL",
                    "m_n_d7_d2" = "m_n_d7_SL",
                    
                    "m_neo_d2" = "m_n_d2",

 
                    "m_n_d2" = "m_neo_d2",)) 

all_crosses <- all_crosses %>%
  mutate(PBbase = recode(PBbase,
                    "adult112" = "h_a_112", 
                    
                    "H_i_120" = "h_i_120", 
                    "H_I.120" = "h_i_120", 
                    "nh120" = "h_i_120", 
                    
                    "H_i_124" = "h_i_124", 
                    "H_I.124" = "h_i_124", 
                    "h_n_124" = "h_i_124",
                    "hi_124" = "h_i_124",
                    

                    "m_neo_d7" = "m_n_d7_SL",
                    "m_n_d7_d2" = "m_n_d7_SL",
                    
                    "m_neo_d2" = "m_n_d2" ) )



```



Now we want to remove the cases where it is same-same
```{r filter sames}

filt_crosses <- all_crosses %>% group_by(imp) %>% filter(PBbase != REFbase) %>% ungroup() %>%
  mutate(cell = ifelse(PBbase == "h_a_112", Leydig, Sertoli))%>%
  filter(PBbase != "m_n_d7_SL" & REFbase != "m_n_d7_SL")%>%
  rename(Mixture = ref)%>%
  select(Mixture, Sertoli, Leydig, cell, PBbase, REFbase, trues, imp, method)


```



combine with Cibersort ones
```{r combine files}

all_testis <- rbind(filt_crosses, cb_testis) %>%
  filter(PBbase != 'm_n_d7') %>%
  filter(PBbase != 'm_n_d7_sl') %>%
  mutate(
    imp = recode(imp, 'noimp' = "No Imp.", 'none' = "No Imp.", 'alra' = 'ALRA', 'saver' = 'SAVER', 'magic' = 'MAGIC'),
    across(c(method, imp, PBbase, REFbase), as.factor),
    across(c(Sertoli, Leydig, trues), as.numeric),
    
    trues = trues/100)

saveRDS(all_testis,"~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/all_testis_wCB.rds")


```


plotting
```{r plots}


df <- all_testis %>% mutate(cell = if_else(cell < 0, 0, cell))

df$imp <- factor(df$imp, levels = c("ALRA", "MAGIC", "No Imp.", "SAVER"))

df <- df %>%
  mutate(
    method = factor(method, levels = c( "CIBERSORTx","DWLS",  "SVR","OLS")),
    imp = factor(imp, levels = c("No Imp.", "ALRA" , "SAVER", "MAGIC" )))


ticks <- c(0.1,0.25,0.50,0.75,0.90)
plots <- ggplot(data=df, aes(x=trues, y=cell, color =REFbase)) +
  geom_point(size=3, shape='diamond', alpha = 0.65) + 
  stat_summary(fun = "mean", geom = "line", size = 2) +  
  geom_abline(slope=1, linetype="dashed") +
  coord_fixed(ratio=1) +  
  scale_x_continuous(breaks=ticks, limits=c(0, 1)) +  
  scale_y_continuous(breaks=ticks, limits=c(0, 1)) +  
  ylab("Predicted Target Cell %") +
  xlab("Actual Target Cell %") +
  scale_color_d3(labels=c("Adult-H", "Infant-120", "Infant-124", "Day 7-M","Day 2-M")) +    
  theme_bw(base_size = 16, base_family = "Myriad Pro") +
  theme(legend.position="bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(color = "Single Cell Reference") +
  guides(color = guide_legend(title.position = "top"))+
  facet_wrap(method~imp, nrow=4) 

plots

ggsave(plots, dpi = 600, filename = "~/OneDrive - UW/data for deconv/Deconvolution output (F)/Testis/plots/cross_compare_testis.png", width = 9, height=12, units = 'in')




```



```{r testis cross-compare mape}


mapes <-  all_testis %>%  
            group_by(method, imp) %>%
            summarise(mape_value = mape(trues, cell)) %>%
            mutate(percent = mape_value*100) %>%
            ungroup()

rmse <-  all_testis %>%  
            group_by(method, imp) %>%
            summarise(rmse_value = rmse(trues, cell)) %>%
            ungroup()



mape_minmax <-  all_testis %>% 
 group_by(method, imp, REFbase) %>%
  summarise(
    mape_value = mape(trues, cell)) %>%
      ungroup()%>%
    group_by(method, imp) %>%
      summarise(
        min=min(mape_value)*100,
        max=max(mape_value)*100)

rmse_minmax <-  all_testis %>% 
 group_by(method, imp, REFbase) %>%
  summarise(
    rmse_value = rmse(trues, cell)) %>%
      ungroup()%>%
    group_by(method, imp) %>%
      summarise(
        min=min(rmse_value),
        max=max(rmse_value))


```
